### 2014-05-02 Knitr/Noise.Rnw
Prepared the document for noise trade article

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage{listings}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
\begin{document}
\title{Noise trader: random walks or forced march?}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk}}
\date{\today}
\maketitle
\begin{abstract}
Noise traders have been characterised as random providers of liquidity, as neutral equivalents of computor-generated random numbers.  hough noise traders may create risk that will deter rational traders from taking advantage of their knowledge of the fundamental value of securities (De-long) and thereby preventing them being driven into extinction as Friedman had asserted (Fridman), their presence  is usually seen as beneficial as they provide liquidity in a neutral fashion.  However, an analysis of noise trading that seeks to measure the weight of noise traders in the market and the strength of their opinin shows that herding is a common feature, highlighting the non-neutral influence that this speculative activity can impose on securities prices. 
\end{abstract}

\section{Introduction}
\section{Literature}
\section{Methods}
\section{Analysis}
\section{Conclusion}
\end{document}